Physics of Tidal Dissipation in Early-Type Stars and White Dwarfs: Hydrodynamical Simulations of Internal Gravity Wave Breaking in Stellar Envelopes

### 2020-02-25 writeups/Su_IGW_break.tex
typo in abstract

\begin{abstract}
    In binaries composed of early-type stars and white dwarfs, the dominant
    tidal process involves the excitation of internal gravity waves (IGWs),
    which propagate towards the stellar surface, and their dissipation via
    nonlinear wave breaking. We perform 2D hydrodynamical simulations of this
    wave breaking process in a stratified, isothermal atmosphere. We find that,
    after an initial transient phase, the dissipation of the IGWs naturally
    generate a sharp critical layer, separating the lower stationary region
    (with no mean flow) and the upper ``synchronized'' region (with the mean
    flow velocity equal to the horizontal wave phase speed). While the critical
    layer is steepened by absorption of these waves, it is simultaneously
    broadened by Kelvin-Helmholtz instabilities such that, in steady state, the
    critical layer width is determined by the Richardson criterion. We study the
    absorption and reflection of incident waves off the critical layer and
    provide analytical formulae describing its long-term evolution. The result
    of this study is important for characterizing the evolution of tidally
    heated white dwarfs and other binary stars.
\end{abstract}

>>>>

\begin{abstract}
    In binaries composed of early-type stars and white dwarfs, the dominant
    tidal process involves the excitation of internal gravity waves (IGWs),
    which propagate towards the stellar surface, and their dissipation via
    nonlinear wave breaking. We perform 2D hydrodynamical simulations of this
    wave breaking process in a stratified, isothermal atmosphere. We find that,
    after an initial transient phase, the dissipation of the IGWs naturally
    generates a sharp critical layer, separating the lower stationary region
    (with no mean flow) and the upper ``synchronized'' region (with the mean
    flow velocity equal to the horizontal wave phase speed). While the critical
    layer is steepened by absorption of these waves, it is simultaneously
    broadened by Kelvin-Helmholtz instabilities such that, in steady state, the
    critical layer width is determined by the Richardson criterion. We study the
    absorption and reflection of incident waves off the critical layer and
    provide analytical formulae describing its long-term evolution. The result
    of this study is important for characterizing the evolution of tidally
    heated white dwarfs and other binary stars.
\end{abstract}
